Tracing the Stellar Halo of Massive Galaxies to 100 kpc at $0.3<z<0.5$ using the Hyper Suprime-Cam

### 2017-04-17 draft/paper1.tex
Intro: 1-D profile

    Following this initial epoch of early formation, feedback from stars and/or AGN 
    (e.g. \citealt{Sijacki2007, Fabian2012}) efficiently quench star formation in 
    massive galaxies and a large fraction of these massive progenitors become quiescent 
    at $z>1$ (e.g. \citealt{Bezanson2009, Kriek2016}). 
    The second phase in the ``two-phase'' formation scenario is driven by 
    non-dissipative processes such as dry mergers with other galaxies 
    (e.g. \citealt{Naab2006, Khochfar2006}). 
    According to the ``two-phase'' scenario, minor mergers drive the growth of massive 
    galaxies at late times and can help explain the significant increase
    of $R_{\mathrm{e}}$ (e.g. \citealt{Newman2012, vdWel2014}) as well as the build-up 
    of the outer stellar halos of ETGs (e.g.  \citealt{Szomoru2012, Patel2013}).        
    {\bf Words here about mass dependence: in the simulations, a larger fraction of the mass 
    is ex-situ at increasingly higher mass. Seen in early work by Naab, Oser, etc, and in the 
    Illustris papers. First prediction to compare with is that the accreted/ex-situ fraction is 
    a function of mass.}

    In addition, according to this two-phase formation scenario, the late time growth of massive ETGs 
    is intrinsically tied to the growth of their host dark matter halos 
    (e.g. \citealt{Leauthaud2012, Behroozi2013, Shankar2013}). 
    Massive ETGs that live at the center of more massive halos should experience 
    a different mass assembly history and should be subject to a larger merger rate 
    compared to ETGs that live in smaller dark matter halos. 
    A second prediction of the two-phase formation scenario is therefore that the 
    structural properties of massive galaxies should depend on their ``environment''

    In order to truly understand the build-up of the outer parts of massive galaxies, very 
    deep data are needed. Massive ETGs have very steep inner stellar density profiles, 
    extremely extended stellar envelopes, and radial-dependent isophotal shapes which 
    make accurate measurements of \mstar{} and $R_{\mathrm{e}}$ very challenging
    (e.g. \citealt{Bernardi2013, DSouza2014}), 
    especially in the presence of shallow imaging data such as the Sloan Digital Sky 
    Survey (SDSS; e.g. \citealt{SDSSDR7, SDSSDR12}).

    In this paper, we take advantage of new high-resolution (median 
    FWHM${\sim} 0.6^{\arcsec}$ in $i$-band) and deep ($i$-band surface brightness 
    limit $> 28.5$\sb) imaging from the Hyper-Suprime Cam (HSC) Subaru Strategic 
    Program \citep[SSP,][]{HSCDR1} to characterize the outer parts of massive galaxies 
    as a function of their properties. 
    The deep imaging depth and excellent seeing conditions of HSC images make them 
    perfect for mapping the \mstar{} distributions of massive galaxies out to very 
    large radii. 
    With the help of the \redm{} cluster catalog (e.g.\ \citealt{Rykoff2014}; 
    \citealt{Rozo2015b}; please see Section~\ref{ssec:redmapper} for details), we 
    select a large sample (${\sim} 15000$) of massive central 
    galaxies with $0.3 < z < 0.5$ using ${\sim} 100$ square degree of data from the 
    HSC wide layer. In this paper, we use this sample to (a)  
    measure profiles for galaxies out to 100 kpc reliably with the HSC data and (b) examine the implications 
    for the stellar mass function when the outer parts of galaxies are accounted for (c) look at 
    trends in outer envelope properties as a function of stellar mass. In a second paper in this series, we 
    will directly show the environmental dependence of the outer envelope mass.
    %The goal of this paper is to compare the \mden{} profiles and structural scaling 
    %relations of massive central galaxies in both low and high mass dark matter halos. 
    %Fig~\ref{fig:sdss_compare} displays a comparison of the imaging quality of massive 
    %ETGs using SDSS versus using the new HSC data.

    \jenny{I suggest removing this here. I do not think it belongs in the Intro:
    Motivated by the ``two-phase'' scenario, and also due to the difficulty in 
    providing appropriate model to describe the \mstar{} distribution of massive 
    galaxy from its center to the very outskirt, we focus on \mstar{} within two 
    different physical apertures in this work:}

@@

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/average_mass_profiles_fsps1_A}
      \caption{
          \textbf{Left}: Median \mden{} profiles in three total stellar mass bins. 
              Thin grey lines show a random subset of individual profiles. 
              The shaded region highlights the region that could be affected by the PSF. 
              Two vertical lines indicate 10 kpc (thin, dotted line) and
              100 kpc (thick, dash line). ~~~ 
          \textbf{Right}: comparison between our \mden{} profiles, previous observations, 
              and simulations. 
              The solid cyan line shows the median profile of massive elliptical 
              galaxies at $z{\sim} 0$ from \citet[][]{Huang2013a}. 
              The red long-dashed line shows the median profile of massive galaxies at 
              $0.25 \leq z < 0.50$ observed by \textit{HST} from \citet[][]{Patel2013}. 
              The purple short-dashed line shows the median radial stellar distributions 
              in massive halos from simulation using the particle tagging method
              (\citealt{Cooper13}).}
      \label{fig:avg_prof}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    Finally, we also compare with the predicted median \mden{} profile of central 
    galaxies in massive halos ($13.5 < \log M_{200,c} < 14.0$) from cosmological 
    simulation where the \mden{} profiles of galaxies are calculated using the 
    particle-tagging technique (e.g. \citealt{Cooper10}).

@@

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_insitu_accretion}
      \caption{
          \song{New figures to illustrate the assumption of in-situ and accreted mass;
                Need to update the caption.}
          Comparison of the median \mden{} profiles of massive galaxies in the 
          two lower \mtot{} bins from Fig~\ref{fig:avg_prof} with other observations 
          and simulations, focusing on the inner region. 
          The format is similar with the right panel of Fig~\ref{fig:avg_prof}.
          We include: 
          1) the median \mden{} profile of inner component from structure 
          decomposition of massive elliptical galaxies at $z{\sim} 0$ from 
          \citet[][Cyan, solid]{Huang2013a}; 
          2) the median \mden{} profile of massive galaxies at $1.0 < z < 1.5$ 
          from \textit{HST} observations in \citet[][Red, dashed]{Patel2013}; 
          3) the median \mden{} profile of the ``in-situ'' stellar components in 
          simulated massive halos from \citet[][Purple, dot-dashed]{Cooper13};
          4) the \mden{} profile of the very massive cD galaxy at $z{\sim} 1.1$ 
          discovered by \citet[][Yellow, dashed]{Liu2013} in the Hubble 
          Ultra-Deep Field.}
      \label{fig:discussion_1}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    We now investigate to what degree our assumption that the inner 10 kpc \mstar{}
    corresponds to an in-situ mass.    
    Fig \ref{fig:discussion_1} compares the median \mden{} profiles of the 
    galaxies in two \mtot{} bins 
    ($11.4\leq$\logmtot{}$<11.6$ and $11.6\leq$\logmtot{}$<11.8$) with 
    (1) the median \mden{} profiles of massive ETGs at $1.0 < z < 1.5$ in
    \citealt{Patel2013} since they are considered as the progenitors of 
    ${\sim} 10^{11.5} M_{\odot}$ ETGs at $z=0$.    
    (2) the inner components of $z{\sim} 0$ ellipticals from 2-D decomposition 
    (\citealt{Huang2013a}).
    As proxy of the in-situ components, \citet{Huang2013b} showed they are 
    structurally similar to the compact ``red nuggets'' at high-$z$. 
    (3) the ``in-situ'' components of simulated central galaxies in massive halos 
    from \citet{Cooper13} (the inner ${\sim} 5$ kpc is quite uncertain due to the 
    resolution).  
    The comparison confirms that \minn{} should be reasonable proxy of the 
    ``in-situ'' mass\footnote{We convert these \mden{} profiles to the same 
    Chabrier IMF; but there are still differences in median \mstar{} and 
    details in the \m2l{} estimates}.  
    More importantly, it shows that the \mden{} profile at 
    $\mathrm{R} > 15$-20 kpc is likely dominated by stars that are not related 
    to the ``in-situ'' processes.

>>>>

    Following this initial epoch of early formation, feedback from stars and/or AGN 
    (e.g. \citealt{Sijacki2007, Fabian2012}) efficiently quench star formation in 
    massive galaxies and a large fraction of these massive progenitors become 
    quiescent at $z>1$ (e.g. \citealt{Bezanson2009, Kriek2016}). 
    The second phase in the ``two-phase'' formation scenario is driven by 
    non-dissipative processes such as dry mergers with other galaxies 
    (e.g. \citealt{Naab2006, Khochfar2006}). 
    \update{
    To explain the observed significant increase of $R_{\mathrm{e}}$ (e.g. 
    \citealt{Newman2012, vdWel2014}) as well as the build-up of the outer stellar 
    halos of ETGs (e.g. \citealt{Szomoru2012, Patel2013}), the ``two-phase'' 
    scenario suggests that minor mergers should dominate the mass assembly during 
    the second phase (e.g. \citealt{Hilz2012, Hilz2013, Oogi2013, Bedorf2013, 
    Laporte2013}).
    More importantly, both numerical simulation (e.g. \citealt{Oser2010}) and 
    semi-analytic model (SAM; e.g. \citealt{LeeYi2013, LeeYi2017}) agree that the 
    mass fraction of the accreted component should increase with total stellar mass, 
    although different models still predict somewhat different values at fixed 
    stellar mass (e.g. \citealt{Lackner2012, Cooper2013, Qu2017}).
    According to the recent results from Illustris \footnote{
    \texttt{http://www.illustris-project.org/}} simulation 
    (\citealt{Vogelsberger2014}, \citealt{Genel2014}), the fraction of the 
    accreted mass increases significantly with total stellar mass at \logms{}$>11.0$, 
    and the accreted component starts to dominate the stellar mass budget 
    ($f_{\mathrm{accreted}}>0.5$) at \logms{}$>11.5$ (\citealt{RodriguezGomez2016}).
    }
%% ------------------------------------------------------------------------------------ %%

    \update{
    At the same time, some of these ``two-phase'' scenario's predictions still have 
    little observational constraint. 
    One-dimensional analysis used to conclude that \ser{} law can reasonably describe  
    the surface brightness (or stellar mass density) profiles of nearby massive 
    elliptical galaxies (e.g. \citealt{Kormendy2009}; except for the most central 
    region), and the \ser{} index seems to increase with total luminosity (or stellar 
    mass; e.g. \citealt{Graham2013}). 
    There was also suggestion that elliptical galaxies can be divided into two groups
    based on how well their 1-D surface brightness profiles can be described by \ser{}
    law (e.g. \citealt{Schombert2015}). 
    Meanwhile, recent two-dimensional analysis points out that the stellar distributions 
    of massive ETGs are better fit by multiple-component models 
    (e.g. \citealt{Huang2013a, Oh2017}). 
    However, it is still not clear whether these 1-D or 2-D results are consistent with 
    the ``two-phase'' scenario. 
    One of the main limitation here is that one needs very deep observations for large 
    sample of massive ETGs to correctly estimate their total stellar mass 
    (e.g. \citealt{Bernardi2013, DSouza2014}) and study the assembly history via 
    their accreted halos (e.g. \citealt{Capaccioli2015, Iodice2016, Iodice2017}).
    Currently, such works are often limited by the shallow images such as the ones from
    the Sloan Digital Sky Survey (SDSS; e.g. \citealt{SDSSDR7, SDSSDR12}) or the small 
    sample size. 
    }

    In this paper, we take advantage of new high-resolution (median 
    FWHM${\sim} 0.6^{\arcsec}$ in $i$-band) and deep ($i$-band surface brightness 
    limit $> 28.5$\sb) imaging from the Hyper-Suprime Cam (HSC) Subaru Strategic 
    Program \citep[SSP,][]{HSCDR1} to characterize the outer parts of massive galaxies 
    as a function of their properties. 
    The deep imaging depth and excellent seeing conditions of HSC images make them 
    perfect for mapping the \mstar{} distributions of massive galaxies out to very 
    large radii. 
    \update{
    With the help of reliable spectroscopic and photometric redshift 
    (e.g. the ones for \redm{} clusters, \citealt{Rykoff2014}; \citealt{Rozo2015b}),
    we select a large sample (${\sim} 7000$) of massive central galaxies at 
    $0.3 < z < 0.5$ using ${\sim} 100$ square degree of data from the 
    HSC wide layer. 
    In this paper, we use this sample to 
    (a) measure individual \mden{} profiles for galaxies out to 100 kpc reliably with 
    the HSC data, 
    (b) examine the implications for the stellar mass function when the outer parts 
    of galaxies are accounted for, and
    (c) look at trends in outer envelope properties as a function of stellar mass. 
    In a second paper in this series, we will directly show the environmental dependence
    of the outer envelope mass (Huang\etal. in prep.)
    }

    \jenny{I suggest removing this here. I do not think it belongs in the Intro:
    Motivated by the ``two-phase'' scenario, and also due to the difficulty in 
    providing appropriate model to describe the \mstar{} distribution of massive 
    galaxy from its center to the very outskirt, we focus on \mstar{} within two 
    different physical apertures in this work:}
    \song{Since the original paper is quite long and complicated, we decide to define
    the two aperture masses in the beginning, and explain why they are important.
    For the new draft, I think it is Ok to do this later, but more comments on this 
    are welcomed!}

@@

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/average_mass_profiles_fsps1_A}
      \caption{
          \textbf{Left}: Median \mden{} profiles in three total stellar mass bins. 
              Thin grey lines show a random subset of individual profiles. 
              The shaded region highlights the region that could be affected by the PSF. 
              Two vertical lines indicate 10 kpc (thin, dotted line) and
              100 kpc (thick, dash line). ~~~ 
          \textbf{Right}: comparison between our \mden{} profiles, previous observations, 
              and simulations. 
              The solid cyan line shows the median profile of massive elliptical 
              galaxies at $z{\sim} 0$ from \citet[][]{Huang2013a}. 
              The red long-dashed line shows the median profile of massive galaxies at 
              $0.25 \leq z < 0.50$ observed by \textit{HST} from \citet[][]{Patel2013}. 
              The purple short-dashed line shows the median radial stellar distributions 
              in massive halos from simulation using the particle tagging method
              (\citealt{Cooper2013}).}
      \label{fig:avg_prof}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    Finally, we also compare with the predicted median \mden{} profile of central 
    galaxies in massive halos ($13.5 < \log M_{200,c} < 14.0$) from cosmological 
    simulation where the \mden{} profiles of galaxies are calculated using the 
    particle-tagging technique (e.g. \citealt{Cooper2010}).

@@

%% ------------------------------------------------------------------------------------ %% 
  \begin{figure*}
      \centering 
      \includegraphics[width=\textwidth]{fig/redbcg_insitu_accretion}
      \caption{
          \song{New figures to illustrate the assumption of in-situ and accreted mass;
                Need to update the caption.}
          Comparison of the median \mden{} profiles of massive galaxies in the 
          two lower \mtot{} bins from Fig~\ref{fig:avg_prof} with other observations 
          and simulations, focusing on the inner region. 
          The format is similar with the right panel of Fig~\ref{fig:avg_prof}.
          We include: 
          1) the median \mden{} profile of inner component from structure 
          decomposition of massive elliptical galaxies at $z{\sim} 0$ from 
          \citet[][Cyan, solid]{Huang2013a}; 
          2) the median \mden{} profile of massive galaxies at $1.0 < z < 1.5$ 
          from \textit{HST} observations in \citet[][Red, dashed]{Patel2013}; 
          3) the median \mden{} profile of the ``in-situ'' stellar components in 
          simulated massive halos from \citet[][Purple, dot-dashed]{Cooper2013};
          4) the \mden{} profile of the very massive cD galaxy at $z{\sim} 1.1$ 
          discovered by \citet[][Yellow, dashed]{Liu2013} in the Hubble 
          Ultra-Deep Field.}
      \label{fig:discussion_1}
  \end{figure*}
%% ------------------------------------------------------------------------------------ %%

@@

    We now investigate to what degree our assumption that the inner 10 kpc \mstar{}
    corresponds to an in-situ mass.    
    Fig \ref{fig:discussion_1} compares the median \mden{} profiles of the 
    galaxies in two \mtot{} bins 
    ($11.4\leq$\logmtot{}$<11.6$ and $11.6\leq$\logmtot{}$<11.8$) with 
    (1) the median \mden{} profiles of massive ETGs at $1.0 < z < 1.5$ in
    \citealt{Patel2013} since they are considered as the progenitors of 
    ${\sim} 10^{11.5} M_{\odot}$ ETGs at $z=0$.    
    (2) the inner components of $z{\sim} 0$ ellipticals from 2-D decomposition 
    (\citealt{Huang2013a}).
    As proxy of the in-situ components, \citet{Huang2013b} showed they are 
    structurally similar to the compact ``red nuggets'' at high-$z$. 
    (3) the ``in-situ'' components of simulated central galaxies in massive halos 
    from \citet{Cooper2013} (the inner ${\sim} 5$ kpc is quite uncertain due to the 
    resolution).  
    The comparison confirms that \minn{} should be reasonable proxy of the 
    ``in-situ'' mass\footnote{We convert these \mden{} profiles to the same 
    Chabrier IMF; but there are still differences in median \mstar{} and 
    details in the \m2l{} estimates}.  
    More importantly, it shows that the \mden{} profile at 
    $\mathrm{R} > 15$-20 kpc is likely dominated by stars that are not related 
    to the ``in-situ'' processes.
